Shaping the Perception of a Robot by Cognitive Priming

### 2016-05-04 main.tex
Proof reading intro

Being such a central ingredient of the psychological ``sauce'' that binds humans to
robots, the field of human-robot interaction (HRI) has long~\fixme{ref?}
attempted to manipulate the anthropomorphic attributes we ascribe to robots.
While modifying the anthropomorphic \emph{design} of robots is conceptually
easy, influencing the underlying psychological phenomenon is less
obvious, although likely more effective.

One way consists in manipulating the context of use of the robot: when social,
entertaining or playful, it elicits more anthropomorphic attributions compared
routine or focused, serious situations (security, rescue, \etc).
\citet{joosse_what_2013} showed, for instance, that when the same robot (a {\sc
nao} humanoid) was used in a different task context (cleaning task \vs tour
guide), users would ascribe different personalities to the robot. Along the same
line, \citet{goetz_cooperation_2002} found that people prefer a serious robot
for serious tasks and a less serious robot for more playful tasks. Also related,
\citet{kaplan_free_2000} discussed how the uselessness of the robot impacts our
tendency to anthropomorphize it: artificial pets like AIBO have no real purpose,
in a sense that they do not provide any kind of service, and this very aspect
may increase people's tendency to anthropomorphize them.

We may hypothesize that these results are a consequence of different
expectations we form about the robot depending on the context: when a machine is
embedded in a social situation, for instance, we expect it to exhibit certain
social skills. However, the experiments conducted so far explore prototypical
opposite situations (routine \vs playful, typically boring \vs typically
intellectual): while they do confirm that the interaction context impacts our
perception of the robot, the stark contrast of the conditions makes it difficult
to finely analyse the psychological determinants that explain the different
perception of the robot.\fixme{references are needed} In particular, existing
techniques to measure anthropomorphism traditionally rely on questionnaires that
may suffer post-hoc reconstruction (we come back on this hereafter).

This article attempts to address this issue with a different experimental design
and a novel methodology to measure \emph{in-the-moment} anthropomorphic attributions.
Our approach is built on a combination of cognitive priming and eye-tracking,
and evidences specific gaze fixations on the head of the robot when the
participants are primed with higher expectations regarding the cognitive skills
of the robot.

>>>>

Being such a central ingredient of the psychological ``sauce'' that binds humans to
robots, the field of human-robot interaction (HRI) has long
attempted to manipulate the anthropomorphic attributes we ascribe to robots.
While modifying the anthropomorphic \emph{design} of robots is conceptually
easy, influencing the underlying psychological phenomenon is less
obvious, although possibly as effective.

One way consists in manipulating the context of use of the robot: when social,
entertaining or playful, it elicits more anthropomorphic attributions compared
routine or focused, serious situations (security, rescue, \etc).
\citet{joosse_what_2013} showed, for instance, that users would ascribe
different personalities to the same robot (a {\sc nao} humanoid) when it was
used in a different task context (cleaning task \vs tour guide). Along the same
line, \citet{goetz_cooperation_2002} found that people prefer a serious robot
for serious tasks and a less serious robot for more playful tasks. Also related,
\citet{kaplan_free_2000} discussed how the uselessness of the robot impacts our
tendency to anthropomorphize it: artificial pets like AIBO have no real purpose,
in a sense that they do not provide any kind of service, and this very aspect
may increase people's tendency to anthropomorphize them.

We may hypothesize that these results are a consequence of different
expectations we form about the robot depending on the context: when a machine is
embedded in a social situation, we expect it to exhibit certain
social skills. The experiments conducted so far only contrast prototypical
\emph{opposite} interaction situations (routine \vs playful, typically boring \vs typically
intellectual): while they do confirm that the interaction context impacts our
perception of the robot, the stark contrast of the conditions makes it difficult
to finely analyse the psychological determinants that explain the different
perception of the robot. In particular, existing
techniques to measure anthropomorphism traditionally rely on questionnaires that
may suffer post-hoc reconstruction (we further discuss this point hereafter).

This article attempts to address this issue with a different experimental design
and a novel methodology to measure \emph{in-the-moment} anthropomorphic attributions.
Our approach is built on a combination of cognitive priming and eye-tracking,
and evidences specific gaze fixations on the head of the robot when the
participants are primed with higher expectations regarding the cognitive skills
of the robot. This methodology leads us next to uncover and evidence the significant
impact of even subtle contextual cues on humans' tendency to ascribe human
traits to machines.
